---
title: "wearable computers"
author: "Theo Livingwell"
date: "October 26, 2017"
output: html_document
---


Load required packages

```{r}
library(ggplot2)
library(dplyr)
library(magrittr)
```

Q1-2. Load, explore and process activity monitoring data 

```{r}
activity_monitors <- read.csv("D:/Document/Supplement/Data_Sci/Repro_Reseach/activity.csv")
summary(activity_monitors)
names(activity_monitors)
head(activity_monitors)
glimpse(activity_monitors)
head(activity_monitors)

activity_monitors$date <- as.Date(activity_monitors$date)
```


What is the mean total number of steps taken per day?

Q1. Calculate the total number of steps taken per day

```{r}
sum_daily_steps <- activity_monitors %>%
  group_by(date) %>%
  filter(!is.na(steps)) %>%
  summarise(total_daily_steps= sum(steps, na.rm=TRUE))
sum_daily_steps
```

Q2. Histogram of total daily steps

```{r}
ggplot(sum_daily_steps, aes(x=total_daily_steps)) + geom_histogram(
  fil="red", binwidth = 1000
) + labs(title = "Total Daily Steps", x="total_daily_steps", y="frequency")
```


Q3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean_daily_steps <- mean(sum_daily_steps$total_daily_steps)

mean_daily_steps

median_daily_steps <- median(sum_daily_steps$total_daily_steps, na.rm = TRUE)

median_daily_steps
```

The mean = 10766.19 and meadian = 10765 are the same suggesting that the data has been properly cleaned 
of outliers 


What is the average daily activity pattern?

Q1. workout average daily steps

```{r}
gaps <- activity_monitors%>%
  group_by (interval)%>%
  filter(!is.na(steps))%>%
  summarise(average_steps = mean(steps, na.rm=TRUE))
gaps
```

Cretaing a time series plot of 5 minutes interval

```{r}
ggplot(gaps, aes(x=interval, y=average_steps)) + 
  geom_line(color="red", size=1) +
  labs(title= "Average Daily Steps", x="interval", y= "Average daily steps")
```


Q2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
gaps[which.max(gaps$average_steps), ]
```

Interval 835 contains 206.1698 average step which is the maximum

Imputing missing values

Q1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(activity_monitors$steps))
```


Q2-3. use mean/median for that day, to fill in NAs. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity_monitors_fill <-activity_monitors
missingValues <- is.na(activity_monitors_fill$steps)
averageInterval <- tapply(activity_monitors_fill$steps, activity_monitors_fill$interval, mean, na.rm=TRUE, simplify = TRUE)
activity_monitors_fill$steps[missingValues] <- averageInterval[as.character(activity_monitors_fill$interval[missingValues])]
names(activity_monitors_fill)

sum(is.na(activity_monitors_fill))      # checking for missing Values, NAs 
```


Reorganise table for better presentation

```{r}
activity_monitors_fill <- activity_monitors_fill[, c("date", "steps", "interval")]
head(activity_monitors_fill, 5)
```

Q4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
total_daily_steps <- activity_monitors_fill%>%
  group_by(date)%>%
  summarise(total_step =  sum(steps, na.rm=TRUE))
total_daily_steps

ggplot(total_daily_steps, aes(x=total_step)) + geom_histogram(fill="red", binwidth = 1000) +
  labs(title = "Histogram of total number of steps per day, includes filled in NAs",  x = "Interval",
       y = "No. of steps")

mean_daily_steps2 <- mean(total_daily_steps$total_step, na.rm = TRUE)
mean_daily_steps2

median_daily_steps2 <- median(total_daily_steps$total_step, na.rm = TRUE)
median_daily_steps2

```

The mean and median are both the same (10766.19) with all the NAs filled in


Are there differences in activity patterns between weekdays and weekends?


Q1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
head(activity_monitors_fill)

```


```{r}
activity_monitors_fill <- activity_monitors_fill %>%
  mutate(weektype = ifelse(weekdays(activity_monitors_fill$date) == "saturday" | weekdays(activity_monitors_fill$date) == "sunday", "weekend", "weekday"))
head(activity_monitors_fill)
```





```{r}
interval_df<- activity_monitors_fill %>% 
  group_by(interval, weektype) %>%
  summarise(average_step2 = mean(steps, na.rm=TRUE))

head(interval_df)
```


Q2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.


```{r}

graph <- ggplot(interval_df, aes(x=interval, y=average_step2, color = weektype)) +
  
  geom_line() +
  labs(title = "Average Steps per Interval Based on Weektype", x = "Interval", y = "Average Number of Steps") +
  facet_wrap(~weektype, ncol = 1, nrow = 2)

print(graph)
```


The step activity is more active early in the day on weekdays. It tends to decrease in activeness in the later parts of the weekdays.


